Paper

### 2022-06-09 Paper.Rmd
minor updates  for tidy

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% filter(label != 'cx') %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
  mutate(chr = factor(chr,levels=c(1:22))) %>% 
  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best') %>% as_tibble %>% 
                                               mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
  rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))

blues = RColorBrewer::brewer.pal(5,'Greys')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label_repel(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('FPR ', round(1-Specificity,2), ' TPR ', round(Sensitivity,2), '\nAUC ',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='Discovery cohort: 50 models', subtitle='18:70 random patient splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
rm(insilico.rocs, insilico.perf, blues, i.thresholds, df, p)
```

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}

tmp = pg.samp %>% dplyr::select(Patient, Status, Endoscopy.Year, PID, Prediction) %>%
  arrange(Patient, Endoscopy.Year, PID)

@@

m = bind_rows(
      get.roc.stat(roc,0.3),
      get.roc.stat(rocNOHGD,0.3)
#      get.roc.stat(rocNOLGD)
      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))

@@

rm(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo, rocNOHGD, rocNOLGD, rocPtM, rocPtA)
```

@@

# discovery + validation
model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
d.n.patient = length(unique(pg.samp$Patient))
d.roc = pROC::roc(Status ~ Prediction, data = pg.samp, ci=T, of='thresholds')

load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
d.coefs = coefs[['0.9']]
d.perf = performance.at.1se[['0.9']]$mean
rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)

@@

sem<-function(x) {
  sd(x)/sqrt(10)
}
# 
# est = info  %>% dplyr::filter(!grepl('SNP', Cohort)) %>% 
#   dplyr::mutate(n.patients = as.numeric(n.patients)) %>%
#   group_by(n.patients, Cohort, Type) %>% 
#   dplyr::summarise_at(vars(-n.samples,-threshold), list(~mean(.,na.rm=T))) %>% 
#   arrange(n.patients) 
# 
# ggplot(est, aes(n.patients, auc)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
# ggplot(est, aes(n.patients, specificity)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
# ggplot(est, aes(n.patients, sensitivity)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
# 
# fit = lm(sensitivity~n.patients, data=est)
# predict(fit, data.frame(n.patients = c(164, 190, 200, 250, 300)), interval = 'confidence' )
# 
# fit = lm((1-specificity)~n.patients, data=est)
# predict(fit, data.frame(n.patients = c(190, 200, 250, 300)), interval = 'confidence' )
# 
# fit = lm(auc~n.patients, data=est)
# predict(fit, data.frame(n.patients = c(190, 200, 250, 300, 325)), interval = 'confidence' )
# 
# fit = lm(n.patients~auc, data=est)
# predict(fit, data.frame(auc = c(0.7, 0.89, 0.92, 0.95)), interval = 'confidence' )

>>>>

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% filter(label != 'cx') %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
  mutate(chr = factor(chr, levels=c(1:22), ordered=T)) %>% 
  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
  left_join(chr.lengths %>% dplyr::select(chrom, chr.length), by=c('chr'='chrom')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best') %>% as_tibble %>% 
                                               mutate(AUC = as.numeric(r$auc), model=r$model, is=as.character(r$is))) ) %>%
  rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))

blues = RColorBrewer::brewer.pal(5,'Greys')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label_repel(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('FPR ', round(1-Specificity,2), ' TPR ', round(Sensitivity,2), '\nAUC ',round(AUC,2))), color='black') +
    scale_color_brewer(palette='Blues') +
    labs(title='Discovery cohort: 50 models', subtitle='18:70 random patient splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
rm(insilico.rocs, insilico.perf, blues, i.thresholds, df, p)
```

@@

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
tmp = pg.samp %>% dplyr::select(Patient, Status, Endoscopy.Year, PID, Prediction) %>%
  arrange(Patient, Endoscopy.Year, PID)

@@

m = bind_rows(
      get.roc.stat(roc,0.3),
#      get.roc.stat(rocNOHGD,0.3)
#      get.roc.stat(rocNOLGD)
      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))

@@

rm(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo, rocPtM, rocPtA)
```

@@

# discovery + validation
model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T)[2], verbose=T)
#load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
d.n.patient = length(unique(pg.samp$Patient))
d.roc = pROC::roc(Status ~ Prediction, data = pg.samp, ci=T, of='thresholds')

load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T), verbose=T)
d.coefs = coefs[['0.9']]
d.perf = performance.at.1se[['0.9']]$mean
rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)

@@

sem<-function(x) {
  sd(x)/sqrt(10)
}
# 
est = info  %>% dplyr::filter(!grepl('SNP', Cohort)) %>%
  dplyr::mutate(n.patients = as.numeric(n.patients)) %>%
  group_by(n.patients, Cohort, Type) %>%
  dplyr::summarise_at(vars(-n.samples,-threshold), list(~mean(.,na.rm=T))) %>%
  arrange(n.patients) %>% ungroup %>%
  dplyr::mutate(idx = row_number())
# 
# ggplot(est, aes(n.patients, auc)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
# ggplot(est, aes(n.patients, specificity)) + geom_point() + geom_smooth(method = 'lm')  + ylim(0.6,1)
# ggplot(est, aes(n.patients, sensitivity)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
# 
fit = lm(sensitivity~n.patients, data=est)
predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )
# 
fit = lm((specificity)~n.patients, data=est)
predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )
# 
fit = lm(auc~n.patients, data=est)
predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )

fit = lm(n.coefs~n.patients, data=est)
autoplot(fit)
predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )

# 
#fit = lm(n.patients~auc, data=est)
#predict(fit, data.frame(auc = c(0.7, 0.89, 0.92, 0.95, 0.97)), interval = 'confidence' )
